Computing Entropies With Nested Sampling

### 2017-07-11 paper/ms.tex
averages -> totals
Appendix references

The idea is to generate a reference particle $\xref$
from the distribution $p(\x)$
whose entropy is required. Then, a Nested Sampling run evolves a set of
particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$,
towards $\xref$ in order to measure the log-probability near
$\xref$ (see Figure~\ref{fig:fig:algorithm}).
Nearness is defined using a distance function
$d(\x; \xref)$, and the number of NS iterations taken to reduce
the distance to below some threshold $L$ provides an unbiased estimate of
\begin{align}
   \depth = -\log \left[ P(d(\x; \xref) < L) \right]
\end{align}
which I call the `depth'. E.g., if $N=10$ Nested Sampling particles are used,
and it takes 100 iterations to reduce the distance to below $L$, then the
depth is 10 nats.

@@

If the threshold $L$ is not small, we might think we are calculating the entropy associated
with the question `what is $x$, to within a tolerance of $\pm L$?'.
This is not quite correct. See the Appendix for a discussion of this
type of question.

@@

The difference in mutual informations between the two observation
schedules is trivial.
The situation may be different if we
had allowed for shorter periods to be possible, as irregular observing
schedules are known to reduce aliasing and ambiguity in the inferred period.
When inferring the period of an oscillating signal, multimodal posterior
distributions for the period are common \citep{gregoryTrimodal, exoplanet}.
The multimodality of the posterior here raises an interesting issue. Is
the question we really want answered ``what is the value of $T$
{\bf precisely}?'', to which the mutual information relates?
Most practicing scientists would not feel particularly informed to learn
that the vast majority of possibilities had been ruled out, if the
posterior still consisted of several widely separated modes!
Perhaps, in some applications, a more appropriate question is
``what is the value of $T$ to within $\pm$ 10\%'', or something along these
lines.

@@

Consider $n=100$ data values, partitioned into two halves, so
$\boldsymbol{y} = \{x_1, ..., x_{50}\} = \{y_1, ..., y_{50}\}$
is the first half of the data and
$\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
is the second half. Let $\ytot$ and $\ztot$ be the averages of the two
subsets of the data:
\begin{align}
\ytot &= \sum_{i=1}^{50} y_i \\
\ztot &= \sum_{i=1}^{50} z_i.
\end{align}
In this section we use the algorithm to answer the questions:
i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$; and
iii) How informative is $\ytot$ about $\ztot$? These are quantified by
$H(\ytot)$, $H(\ztot)$, and $I(\ytot; \ztot)$ respectively. This can
be done despite the fact that $\alpha$ is unknown. The joint distribution
for $\ln \ytot$ and $\ln \ztot$, whose properties I am to obtain,
is shown in Figure~\ref{fig:pareto}.

@@

\appendix
\section{Precisional questions}

@@

\appendix
\section{Software}

>>>>

The idea is to generate a reference particle $\xref$
from the distribution $p(\x)$
whose entropy is required. Then, a Nested Sampling run evolves a set of
particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$,
towards $\xref$ in order to measure the log-probability near
$\xref$ (see Figure~\ref{fig:algorithm}).
Nearness is defined using a distance function
$d(\x; \xref)$, and the number of NS iterations taken to reduce
the distance to below some threshold $L$ provides an unbiased estimate of
\begin{align}
   \depth = -\log \left[ P(d(\x; \xref) < L) \right]
\end{align}
which I call the `depth'. E.g., if $N=10$ Nested Sampling particles are used,
and it takes 100 iterations to reduce the distance to below $L$, then the
depth is 10 nats.

@@

If the threshold $L$ is not small, we might think we are calculating the entropy associated
with the question `what is $x$, to within a tolerance of $\pm L$?'.
This is not quite correct. See appendix~\ref{sec:precisional}
for a discussion of this
type of question.

@@

The difference in mutual informations between the two observation
schedules is trivial.
The situation may be different if we
had allowed for shorter periods to be possible, as irregular observing
schedules are known to reduce aliasing and ambiguity in the inferred period.
When inferring the period of an oscillating signal, multimodal posterior
distributions for the period are common \citep{gregoryTrimodal, exoplanet}.
The multimodality of the posterior here raises an interesting issue. Is
the question we really want answered ``what is the value of $T$
{\bf precisely}?'', to which the mutual information relates?
Most practicing scientists would not feel particularly informed to learn
that the vast majority of possibilities had been ruled out, if the
posterior still consisted of several widely separated modes!
Perhaps, in some applications, a more appropriate question is
``what is the value of $T$ to within $\pm$ 10\%'', or something along these
lines. See Appendix~\ref{sec:precisional} for more information about this
issue.

@@

Consider $n=100$ data values, partitioned into two halves, so
$\boldsymbol{y} = \{x_1, ..., x_{50}\} = \{y_1, ..., y_{50}\}$
is the first half of the data and
$\boldsymbol{z} = \{x_{51}, ..., x_{100}\} = \{z_1, ..., z_{50}\}$
is the second half. Let $\ytot$ and $\ztot$ be the totals of the two
subsets of the data:
\begin{align}
\ytot &= \sum_{i=1}^{50} y_i \\
\ztot &= \sum_{i=1}^{50} z_i.
\end{align}
In this section we use the algorithm to answer the questions:
i) How uncertain is $\ytot$?; ii) How uncertain is $\ztot$; and
iii) How informative is $\ytot$ about $\ztot$? These are quantified by
$H(\ytot)$, $H(\ztot)$, and $I(\ytot; \ztot)$ respectively. This can
be done despite the fact that $\alpha$ is unknown. The joint distribution
for $\ln \ytot$ and $\ln \ztot$, whose properties I am to obtain,
is shown in Figure~\ref{fig:pareto}.

@@

\appendix
\section{Precisional questions}
\label{sec:precisional}

@@

\appendix
\section{Software}
\label{sec:software}
